---
title: "Developing Data Products Shiny Application Project"
author: "Kyle Ward"
date: "March 28, 2019"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
library(rsconnect)
library(leaflet)
library(plotly)
library(tidyverse)
library(widgetframe)
library(Cairo)

env1 = new.env()

state_table_by_actyr <- readRDS("C:/Users/kward/DataScience/Fannie Mae ETL/Hackathon/Data_Products_Project/state_table_by_actyr.rds")
us_geo <- readRDS("C:/Users/kward/DataScience/Fannie Mae ETL/Hackathon/Data_Products_Project/us_geo.rds")
b <- readRDS("C:/Users/kward/DataScience/Fannie Mae ETL/Hackathon/Data_Products_Project/orig_amt_by_yr.rds")





```



## Fannie Mae Overview

- Fannie Mae is a US government-sponsored enterprise responsible for supporting the secondary mortgage market
- Fannie Mae buys mortgage loans and mortgages in order to securitize the loans
- A web-based interactive application was created to provide loan performance analytics based on criteria selected by the end user
- Data Source 
    - Fannie Mae releases a sample of their loan portfolio's performance data in order to support public understanding
    - The sample used in this project consists of 30-year and less, fully amortizing, full documentation, single-family, conventional fixed-rate mortgages

  
## Loan Acquisition Profile
```{r acquisition plot, echo=FALSE}
ay <- list(
  tickfont = list(color = "black"),
  overlaying = "y",
  side = "right",
  title = "number of loans"
)

p3 <- plot_ly(b,x=~vinyr, y=~total_orig_amt, name="UPB" ,type='bar' ) %>% 
        layout(title="Loan Origination UPB by cohort",yaxis=list(title="Loan Origination UPB ($)"),yaxis2=ay,
         xaxis=list(title="Cohort"))
 

p3

```

- The loan acquisition profile portrays the value of Fannie Mae loans originated from 2000 to 2017

## Loan Performance Application Objective

- The Shiny application was developed to provide insights into the peformance of the loan portfolio over time
- The application will allow the user to visualize the default rate for each state to identify geographicloan performance trends
- The geospatial app can be found on github by selecting the following link: <link>



## Interactive GeoSpatial App
```{r, echo=FALSE}
shinyApp(
ui=fluidPage(
  titlePanel(p("Default Rate by State by Year", style = "color:#3474A7")),
  includeMarkdown('html/geospatial_description.md'),
  sidebarLayout(
      sidebarPanel(
        selectInput(inputId = "activity_year_selected", label = "Select  activity year",
                    choices = 2000:2018)),
      mainPanel(
        leafletOutput(outputId = "map")
        )
  )
),
server= function(input, output) {
    output$map = renderLeaflet({
      filtered_merge <-  state_table_by_actyr %>% 
        dplyr::ungroup() %>% 
        filter(actyr == input$activity_year_selected) %>%
        tmaptools::append_data(
          shp = us_geo, 
          data =  ., 
          key.shp = "STUSPS", 
          key.data = "state", 
          ignore.duplicates = TRUE)
        
      bins <- c(0, .01, .03, .06, .2, .25, .3, .35)
      pal <- colorBin("YlOrRd", domain = filtered_merge$default_rate, bins = bins)
      
      labels <- sprintf(
        "State: %s <br> default rate: %g",
         filtered_merge$STUSPS, filtered_merge$default_rate) %>% 
        lapply(htmltools::HTML)
      
      leaflet(filtered_merge) %>%
        setView(-96, 37.8, 4) %>%
        addTiles() %>%
        addPolygons(
          fillColor = ~pal(default_rate),
          weight = 2,
          opacity = 1,
          color = "white",
          dashArray = "3",
          fillOpacity = 0.7,
          highlight = highlightOptions(
            weight = 5,
            color = "#666",
            dashArray = "",
            fillOpacity = 0.7,
            bringToFront = TRUE),
          label = labels,
          labelOptions = labelOptions(
            style = list("font-weight" = "normal", padding = "3px 8px"),
            textsize = "15px",
            direction = "auto")) %>%
        addLegend(pal = pal, values = ~default_rate, opacity = 0.7, title = "Default Rate",
                  position = "bottomright")})
},
options = list(height = 200)
  )   

```



## Thank You!




